HMM Model development

### 2020-05-14 dev/hmm_model_devel.Rmd
Including noisy states

```{r}
model <- stan_model(here::here("hmm.stan"))
```

```{r}
adapt_delta <- 0.8

data <- hmm_simulator(10, 20, 2, use_noisy_states = TRUE)

if(inherits(model, "stanmodel")) {
  fit <- sampling(model, data = data$observed, control = list(adapt_delta = adapt_delta))
} else {
  cmdstan_fit <- model$sample(data = data$observed, adapt_delta = adapt_delta)
  fit <- rstan::read_stan_csv(cmdstan_fit$output_files())
}
evaluation_summary(fit, data$true)

@@

```{r}
sbc_res <- sbc(model, generator = function() {hmm_simulator(10, 20, 2, use_noisy_states = FALSE) }, N_steps = 100, control = list(adapt_delta = adapt_delta))

@@

sbc_res$params %>% filter(grepl("mu", param_name)) %>% plot_sbc_params()
sbc_res$params %>% filter(grepl("sensitivity", param_name)) %>% plot_sbc_params()
sbc_res$params %>% filter(grepl("other_observations_probs", param_name)) %>% plot_sbc_params()

@@

```{r}
data <- hmm_simulator(5, 10, 2, use_noisy_states = FALSE)
cc <- make_stancode_hmm(data$observed_structured$formula, serie_data = data$observed_structured$serie_data, rate_data = data$observed_structured$rate_data, state_data = data$observed_structured$state_data, prior= data$observed_structured$prior)

model2 <- stan_model(model_code = cc)

prep_data <- make_data_hmm(data$observed_structured$formula, serie_data = data$observed_structured$serie_data, rate_data = data$observed_structured$rate_data, state_data = data$observed_structured$state_data)

all_data <- make_standata_hmm(data$observed_structured$formula, serie_data = data$observed_structured$serie_data, rate_data = data$observed_structured$rate_data, state_data = data$observed_structured$state_data)

fit_raw <- sampling(model2, data = all_data)
#TODO - compactification is necessary!
```

>>>>

```{r}
adapt_delta <- 0.9

data <- hmm_simulator(30, 20, 2, use_noisy_states = TRUE)

single_code <- make_stancode_hmm(data$observed_structured$formula, serie_data = data$observed_structured$serie_data, rate_data = data$observed_structured$rate_data, state_data = data$observed_structured$state_data, prior= data$observed_structured$prior)

model_single <- stan_model(model_code = single_code)

prepdata <- make_data_hmm(data$observed_structured$formula, serie_data = data$observed_structured$serie_data, rate_data = data$observed_structured$rate_data, state_data = data$observed_structured$state_data, sensitivity_low_bound = data$observed$sensitivity_low_bound)

standata <- make_standata_hmm(data$observed_structured$formula, serie_data = data$observed_structured$serie_data, rate_data = data$observed_structured$rate_data, state_data = data$observed_structured$state_data, sensitivity_low_bound = data$observed$sensitivity_low_bound)

if(inherits(model_single, "stanmodel")) {
  fit <- sampling(model_single, data = standata, control = list(adapt_delta = adapt_delta))
} else {
  cmdstan_fit <- model_single$sample(data = standata, adapt_delta = adapt_delta)
  fit <- rstan::read_stan_csv(cmdstan_fit$output_files())
}
evaluation_summary(fit, data$true)

@@

```{r}
noisy_states_in_sbc <- FALSE
sbc_generator <- function() {
  data <- hmm_simulator(50, 20, 3, use_noisy_states = noisy_states_in_sbc)
  standata <- make_standata_hmm(data$observed_structured$formula, serie_data = data$observed_structured$serie_data, rate_data = data$observed_structured$rate_data, state_data = data$observed_structured$state_data, sensitivity_low_bound = data$observed$sensitivity_low_bound)

  list(observed = standata,
       true = data$true)
}

sbc_res <- sbc(model_single, generator = sbc_generator, N_steps = 200, control = list(adapt_delta = adapt_delta))

@@

sbc_res$params %>% filter(grepl("b|sd", param_name)) %>% plot_sbc_params()
sbc_res$params %>% filter(grepl("r_1", param_name)) %>% plot_sbc_params()
sbc_res$params %>% filter(grepl("r_2", param_name)) %>% plot_sbc_params()
if(noisy_states_in_sbc) {
  sbc_res$params %>% filter(grepl("sensitivity", param_name)) %>% plot_sbc_params()
  sbc_res$params %>% filter(grepl("other_observations_probs", param_name)) %>% plot_sbc_params()
}

@@

```{r}
```
